Variational Monte Carlo on bosonic systems

### 2018-04-03 project_1/article/article.tex
Add section on stability

            Scanning the entire space of variational parameters can be a
            somewhat despairing task. To aid us in the search we therefore
            implement the method of gradient descent\footnote{Also known as
            steepest descent, easily confused with the steepest descent method
            for integation.}. In any optimum we generally have that the
            derivative should equal zero. In our case, we want to find the
            minimum of the expected local energy with respect to some
            variational parameter $\alpha$, so that
            \begin{equation}
                \nabla_\alpha \expval{E_L(\mathbf{R}; \alpha)} = 0.
            \end{equation}
            The foundation for  the method of gradient descent lies in the fact
            that some analytic \footnote{We use the term analytic quite
            recklessly.} function $F(\alpha)$ decreases fastest in the direction
            of $-\nabla_\alpha F(\alpha)$. This means that we could eventually
            get to an ensuing minimum by trailing along the pathway of the
            following recursive relation,
            \begin{equation}
                \alpha_{k+1} = \alpha_k - \gamma \nabla_{\alpha} F(\alpha_k).
            \end{equation}
            If $\gamma$ is small enough we should have $f(\alpha_k) \geq
            F(\alpha_{k+1})$ for all $k$ and hopefully we will experience
            convergence to the desired minimum.

            For our special case we only have one variational parameter,
            $\alpha$.  The expression for the derivative is,
            \begin{equation}
                \nabla_{\alpha}\expval{E_L}
                =
                2\left(
                    \expval{E_L\frac{1}{\Psi_T}\nabla_{\alpha}\Psi_T}
                    - \expval{\frac{1}{\Psi_T} \nabla_{\alpha}\Psi_T}\expval{E_L}
                \right).
                \label{eq:variational_parameter_gradient}
            \end{equation}

            %TODO: INSERT DERIVATION OF LOCAL ENERGY??

>>>>

            Scanning the entire space of variational parameters can be a
            somewhat despairing task. To aid us in the search we therefore
            implement the method of gradient descent\footnote{Also known as
            steepest descent, easily confused with the steepest descent method
            for integation.}. In any optimum we generally have that the
            derivative should equal zero. In our case, we want to find the
            minimum of the expected local energy with respect to some
            variational parameter $\alpha$, so that
            \begin{equation}
                \nabla_\alpha \expval{E_L(\mathbf{R}; \alpha)} = 0.
            \end{equation}
            The foundation for  the method of gradient descent lies in the fact
            that some analytic \footnote{We use the term analytic quite
            recklessly.} function $\vf{F}(\alpha)$ decreases fastest in the direction
            of $-\nabla_\alpha\cdot\vf{F}(\alpha)$. This means that we could eventually
            get to an ensuing minimum by trailing along the pathway of the
            following recursive relation,
            \begin{equation}
                \vfg{\alpha}_{k+1}
                = \vfg{\alpha}_k - \gamma \nabla_{\alpha}\cdot \vf{F}(\alpha_k).
                \label{eq:parameter_difference}
            \end{equation}
            If $\gamma$ is small enough we should have $\vf{F}(\alpha_k) \geq
            \vf{F}(\alpha_{k+1})$ for all $k$ and hopefully we will experience
            convergence to the desired minimum.

            For our special case we only have one variational parameter,
            $\alpha$. The expression for the derivative is,
            \begin{align}
                \nabla_{\alpha}\expval{E_L}
                =
                2\left(
                    \expval{E_L\frac{1}{\Psi_T}\nabla_{\alpha}\Psi_T}
                    - \expval{\frac{1}{\Psi_T} \nabla_{\alpha}\Psi_T}
                    \expval{E_L}
                \right).
                \label{eq:variational_parameter_gradient}
            \end{align}
            To evaluate this expression we sample the local energy, the
            parameter gradient of the wavefunction, and the combined expression
            of these at each Monte Carlo cycle.

            \subsubsection{Stability of the algorithm}
                When employing the method of gradient descent to locate a
                minimum value for the variational parameters we expect some
                issues due to stability of the Monte Carlo methods and the
                steepnes of the expected energy curves. If our guess of the
                location of the minimum is far off we can expect the method to
                diverge as the gradient will become huge. This will yield a
                new value for $\vfg{\alpha}$ in
                \autoref{eq:parameter_difference} which will be further away and
                might result in a local minimum ``trapping'' the method. To
                prevent this from happening we have added a limit to the allowed
                change in $\vfg{\alpha}$. If the change in $\vfg{\alpha}$ is
                larger than a set order of magnitude we keep dividing
                $\vfg{\alpha}$ with this number until this criteria is
                fulfilled.
